---
title: "China hits back, calculated from China data, not from the US data"
author: "Mitsuo Shiota"
date: "2019-06-18"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Updated: `r Sys.Date()`

## Summary

Since [I guessed that Chinese calculated their retaliation values on HS 6 digit codes by ignoring 8 digit codes based on the data from the US customs](China-hits-back.md), I have looked for the data from China customs, and found [UN Comtrade](https://comtrade.un.org/) has HS 6 digit code data reported by China.


## Libraries

As usual, I load tidyverse package.

```{r r libraries, include=FALSE}
library(tidyverse)

```

I prepare some functions.

CORRECTION: I mistakenly let text2df function drop 0s in the right. This error makes matching failures, and reduces the calculated retaliation values. For the record, I leave this error uncorrected. For the corrected version, look for [a new .Rmd file](https://github.com/mitsuoxv/us-tariffs-on-china/blob/master/China-hits-back3.Rmd) and [a new .md file](https://github.com/mitsuoxv/us-tariffs-on-china/blob/master/China-hits-back3.md).

```{r functions}
# download to temporary pdf file, and scratch text from pdf
url2text <- function(url) {
  tf <- tempfile(fileext = ".pdf")
  
  httr::GET(url, httr::write_disk(tf))
  
  pdftools::pdf_text(tf)
}

# scratch pattern from text
text2df <- function(text, pattern, tariff) {
  tariff_list <- text %>% 
    str_extract_all(pattern) %>% 
    unlist() %>% 
    str_replace("0+$", "")
  
  tibble(hs = tariff_list, tariff = tariff)
}

```

## Get retaliation tariff lists in HS 8 digit codes

I download Excel and PDF files from the Chinese government sites. I scratch HS 8 digit codes from those files. And I get the tariff list of each iteration or tranche like "3b", "34b", "16b" and "60b".

The difference from [the original study](China-hits-back.md) in this part is that I don't omit 0s in the codes in the right.

```{r df_list, echo=FALSE, results="hide", cache=FALSE}
# Retariation to Iron and Aluminium
# http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201804/t20180401_2857769.html
# url <- "http://http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201804/P020180401856022545193.xlsx"
# httr::GET(url, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))

res <- readxl::read_excel("data/P020180401856022545193.xlsx", col_names = TRUE, skip = 4)

names(res) <- c("item_number", "hs", "description", "rates")

df_list_3b <- res %>% 
  select(hs)

nrow(df_list_3b) # 128

df_list_3b$tariff <- "3b"

# Tariff lists in HS 8 digit codes
hs <- "\\d{8}"

# US first tranche 34 billion dollars, 25 percent, effective on July 6, 2018
# http://www.mofcom.gov.cn/article/ae/ai/201806/20180602756389.shtml
# Retariation by China is the same amount
text <- url2text("http://images.mofcom.gov.cn/www/201806/20180616015345014.pdf")

df_list_34b <- text2df(text, hs, "34b")

nrow(df_list_34b) # 545


# US second tranche 16 billion dollars, 25 percent, August 23, 2018
# Retariation by China is the same amount
text <- url2text("http://images.mofcom.gov.cn/www/201806/20180616015405568.pdf")

df_list_16b <- text2df(text, hs, "16b")

nrow(df_list_16b) # 114

# US 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/t20180803_2980950.html
# Group 1, 25 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711628159425.pdf")

df_list_60b_1 <- text2df(text, hs, "60b")

nrow(df_list_60b_1) # 2494, different from 2493

df_list_60b_1 <- df_list_60b_1 %>%
  unique()

nrow(df_list_60b_1) # 2493

# 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# Group 2, 20 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711628507212.pdf")

df_list_60b_2 <- text2df(text, hs, "60b")

nrow(df_list_60b_2) # 1078

# 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# Group 3, 10 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711628844240.pdf")

df_list_60b_3 <- text2df(text, hs, "60b")

nrow(df_list_60b_3) # 974

# 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# Group 4, 5 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711629144210.pdf")

df_list_60b_4 <- text2df(text, hs, "60b")

nrow(df_list_60b_4) # 662

# Combine 60B
df_list_60b <- df_list_60b_1 %>% 
  bind_rows(df_list_60b_2) %>% 
  bind_rows(df_list_60b_3) %>% 
  bind_rows(df_list_60b_4)

```

## Cut 8 to 6 digit codes

I cut HS codes from 8 to 6 digits.

```{r list_6, echo=FALSE}
df_list_3b_6 <- df_list_3b %>% 
  mutate(
    hs = hs %>% str_sub(end = 6L)
  ) %>% 
  unique()

df_list_34b_6 <- df_list_34b %>% 
  mutate(
    hs = hs %>% str_sub(end = 6L)
  ) %>% 
  unique()

df_list_16b_6 <- df_list_16b %>% 
  mutate(
    hs = hs %>% str_sub(end = 6L)
  ) %>% 
  unique()

df_list_60b_6 <- df_list_60b %>% 
  mutate(
    hs = hs %>% str_sub(end = 6L)
  ) %>% 
  unique()

```

## Get China import from US data in 2017 in HS 6 digit codes

I get the id's of both China and the US by basically following [Using the UN Comtrade data API with R](https://comtrade.un.org/Data/Doc/api/ex/r) page.

```{r get_country_id, cache=FALSE}
string <- "http://comtrade.un.org/data/cache/partnerAreas.json"

reporters <- jsonlite::fromJSON(string)

reporters$results %>% 
  filter(text %in% c("China", "USA"))

```
I make query to extract China import from the US on HS 6 digit in 2017.

```{r get 6 digit data, cache=FALSE}
q_list <- list(
  max=50000,
  type="C",
  freq="A",
  px="HS",
  ps=2017,
  r=156,
  p=842,
  rg=1,
  cc="AG6",
  fmt="json"
)

response <- httr::GET(
  url = "http://comtrade.un.org/api/get",
  query = q_list
)

res <- httr::content(response)

df2017_6 <- res$dataset %>% 
  map_dfr(~ list(hs = .x$cmdCode,
                 value = .x$TradeValue))

total <- df2017_6 %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000 

```
China imports from US in 2017 amount to `r round(total, 0)` billion dollars in total. I have got data by HS 6 digit codes in "df2017_6", a data frame of 2 columns of "hs" and "value."


```{r value_6, echo=FALSE}
value_3b_6 <- df2017_6 %>% 
  semi_join(df_list_3b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000
  
value_34b_6 <- df2017_6 %>% 
  semi_join(df_list_34b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

value_16b_6 <- df2017_6 %>% 
  semi_join(df_list_16b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

value_60b_6 <- df2017_6 %>% 
  semi_join(df_list_60b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

```

The results in "3b", "34b", "16b" and "60b" are `r round(value_3b_6, 2)`, `r round(value_34b_6, 2)`, `r round(value_16b_6, 2)` and `r round(value_60b_6, 2)` billion dollars respectively. Ratios to the Chinese claims are `r round(value_3b_6 / 3, 2)`, `r round(value_34b_6 / 34, 2)`, `r round(value_16b_6 / 16, 2)` and `r round(value_60b_6 / 60, 2)`.

CORRECTION: These numbers are miscaluculated due to an error in text2df function, and are less than correct numbers. For correct numbers, please refer to [a new page](https://github.com/mitsuoxv/us-tariffs-on-china/blob/master/China-hits-back3.md).

EOL
