---
title: "China hits back, half the size of its claim"
author: "Mitsuo Shiota"
date: "2019-04-24"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Updated: `r Sys.Date()`

## Summary

After [I confirmed the USTR's claims on how much imports from China are affected by the tariffs they impose](README.md), I now turn to the Chinese claims on how much imports from the United States are affected by the retaliation tariffs they impose.

If my calculations using [Census Bureau U.S. International Trade Data](https://www.census.gov/foreign-trade/data/) are right, the Chinese hit-backs are half the size of their claims. My guess is that they calculated the size based on HTS 6 digit codes, while they actually impose tariffs based on both 6 and 8 digit codes.

Even if retaliation size is less than claimed, US exports are being hurt.

## Libraries and functions

I use the same packages I used in [README.md](README.md).

```{r r libraries, include=FALSE}
library(tidyverse)

```
I prepare some functions.

```{r functions, echo=FALSE}
# download to temporary pdf file, and scratch text from pdf
url2text <- function(url) {
  tf <- tempfile(fileext = ".pdf")
  
  httr::GET(url, httr::write_disk(tf))
  
  pdftools::pdf_text(tf)
}

# scratch pattern from text
text2df <- function(text, pattern, tariff) {
  tariff_list <- text %>% 
    str_extract_all(pattern) %>% 
    unlist() %>% 
    str_replace("0+$", "")
  
  tibble(hs = tariff_list, tariff = tariff)
}

# name an unnamed list like V1, V2, V3 ...
name_list <- function(list) {
  names(list) <- paste0("V", seq_along(list))
  
  list
}

# transform response list to df
res2df <- function(response) {
  res_content <- httr::content(response)
  
  df <- res_content[-1] %>% 
    map_dfr(name_list)
  
  names(df) <- unlist(res_content[[1]])
  
  df
}

# get country export data by year and by specified HTS digits (1-10)
export_to_country <- function(country, hs_digits, year) {
  try(if(hs_digits < 1 | hs_digits > 10) stop("hs_digits must be between 1 and 10"))
  
  digits <- if_else(hs_digits <=2, 2,
                    if_else(hs_digits <= 4, 4,
                            if_else(hs_digits <= 6, 6, 10)
                    )
  )
  
  response <- httr::GET(
    url = "https://api.census.gov/data/timeseries/intltrade/exports/hs",
    query = list(
      get="ALL_VAL_MO,E_COMMODITY",
      time=year,
      CTY_CODE=country,
      COMM_LVL=str_c("HS", digits),
      key=keyring::key_get("census")
    )
  )
  
  df <- res2df(response)
  
  df$ALL_VAL_MO <- as.numeric(df$ALL_VAL_MO)
  
  df$time <- as.Date(paste0(df$time, "-01"), "%Y-%m-%d")
  
  df %>% 
    rename(
      value = ALL_VAL_MO,
      hs = E_COMMODITY
    ) %>% 
    mutate(
      hs = str_sub(hs, end = hs_digits) # cut codes
    ) %>% 
    # summarize because cutting codes may have created duplication
    group_by(time, hs) %>% 
    summarize(value = sum(value)) %>% 
    ungroup() %>% 
    select(time, hs, value) %>% 
    arrange(hs, time)
}

```
## Get tariff lists from Excel and PDF files

I download Excel and PDF files from the Chinese government sites. I scratch HTS 8 digit codes from those files, and omit 0s in the codes in the right, as I guess 0s in the right means inclusion of any numbers. And I get the tariff list of each iteration or tranche like "3b", "34b", "16b" and "60b".

```{r df_list, echo=FALSE, results="hide", cache=FALSE}
# Retariation to Iron and Aluminium
# http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201804/t20180401_2857769.html
# url <- "http://http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201804/P020180401856022545193.xlsx"
# httr::GET(url, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))

res <- readxl::read_excel("data/P020180401856022545193.xlsx", col_names = TRUE, skip = 4)

names(res) <- c("item_number", "hs", "description", "rates")

df_list_3b <- res %>% 
  select(hs)

nrow(df_list_3b) # 128

df_list_3b$tariff <- "3b"

df_list_3b$hs <- df_list_3b$hs %>% 
  str_replace("0+$", "")

# Tariff lists in HTS 8 digit codes
# HTS (Harmonized Tariff Schedule of the United States)
hts <- "\\d{8}"


# US first tranche 34 billion dollars, 25 percent, effective on July 6, 2018
# http://www.mofcom.gov.cn/article/ae/ai/201806/20180602756389.shtml
# Retariation by China is the same amount
text <- url2text("http://images.mofcom.gov.cn/www/201806/20180616015345014.pdf")

df_list_34b <- text2df(text, hts, "34b")

nrow(df_list_34b) # 545


# US second tranche 16 billion dollars, 25 percent, August 23, 2018
# Retariation by China is the same amount
text <- url2text("http://images.mofcom.gov.cn/www/201806/20180616015405568.pdf")

df_list_16b <- text2df(text, hts, "16b")

nrow(df_list_16b) # 114

# US 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/t20180803_2980950.html
# Group 1, 25 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711628159425.pdf")

df_list_60b_1 <- text2df(text, hts, "60b")

nrow(df_list_60b_1) # 2494, different from 2493

df_list_60b_1 <- df_list_60b_1 %>%
  unique()

nrow(df_list_60b_1) # 2493

# 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# Group 2, 20 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711628507212.pdf")

df_list_60b_2 <- text2df(text, hts, "60b")

nrow(df_list_60b_2) # 1078

# 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# Group 3, 10 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711628844240.pdf")

df_list_60b_3 <- text2df(text, hts, "60b")

nrow(df_list_60b_3) # 974

# 200 billion dollars, 10 percent, September 24, 2018
# Retariation by China is 60 billion dollars in 4 groups
# Group 4, 5 percent
text <- url2text("http://gss.mof.gov.cn/zhengwuxinxi/zhengcefabu/201808/P020180803711629144210.pdf")

df_list_60b_4 <- text2df(text, hts, "60b")

nrow(df_list_60b_4) # 662

# Combine 60B
df_list_60b <- df_list_60b_1 %>% 
  bind_rows(df_list_60b_2) %>% 
  bind_rows(df_list_60b_3) %>% 
  bind_rows(df_list_60b_4)

```

## Get US export to China data in 2017

```{r API_Key_input, include=FALSE}
keyring::key_set("census")

```

```{r get 8 digit data, cache=FALSE}
# Get export from China data
# https://api.census.gov/data/timeseries/intltrade/exports/hs/variables.html

# country code, China 5700
# https://www.census.gov/foreign-trade/schedules/c/countryname.html

df2017 <- export_to_country(5700, 8, 2017)

total <- df2017 %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000 

```
US exports to China in 2017 amount to `r round(total, 0)` billion dollars in total. I have got data by HTS 8 digit codes in "df2017", a data frame of 3 columns of "time", "hs" and "value."

## US export values don't add up to the Chinese claims

I prepare function to forward match 8 digit codes to less than 8 digit codes, in order to make the look-up table of 2 columns, "hs" of unique HTS 8 digit codes and "tariff" of tariff category like "34b".

```{r lookup_hs}
# add tariff category to df2017
lookup_hs <- tibble(hs = df2017$hs %>% unique())

# function
find_tariff <- function(hs, df) {
  patterns <- str_c("^", df$hs)

  result <- df[str_detect(hs, patterns), ] %>% 
    select(tariff) %>% 
    as.character()

  result
}

```
As there are duplication of HTS codes among "tariff", I make one look-up table for each "tariff", left join it to the US export data, and calculate sum of "value" by each "tariff".

```{r 3b, echo=FALSE, cache=FALSE}
lookup_hs_3b <- lookup_hs %>% 
  mutate(tariff = map_chr(hs, find_tariff, df_list_3b))

df2017_3b <- df2017 %>% 
  left_join(lookup_hs_3b, by = "hs")

df_3b <- df2017_3b %>% 
  group_by(tariff) %>% 
  summarize(value = sum(value))

df_3b

value_3b <- df_3b %>% 
  filter(tariff != "character(0)") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

```
US export value in 3b list is `r round(value_3b, 2)` billion dollars, and its ratio to the Chinese claim is `r round(value_3b / 3, 2)`.

```{r 34b, echo=FALSE, cache=FALSE}
lookup_hs_34b <- lookup_hs %>% 
  mutate(tariff = map_chr(hs, find_tariff, df_list_34b))

df2017_34b <- df2017 %>% 
  left_join(lookup_hs_34b, by = "hs")

df_34b <- df2017_34b %>% 
  group_by(tariff) %>% 
  summarize(value = sum(value))

df_34b

value_34b <- df_34b %>% 
  filter(tariff != "character(0)") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

```
US export value in 34b list is `r round(value_34b, 2)` billion dollars, and its ratio to the Chinese claim is `r round(value_34b / 34, 2)`.

```{r 16b, echo=FALSE, cache=FALSE}
lookup_hs_16b <- lookup_hs %>% 
  mutate(tariff = map_chr(hs, find_tariff, df_list_16b))

df2017_16b <- df2017 %>% 
  left_join(lookup_hs_16b, by = "hs")

df_16b <- df2017_16b %>% 
  group_by(tariff) %>% 
  summarize(value = sum(value))

df_16b

value_16b <- df_16b %>% 
  filter(tariff != "character(0)") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

```
US export value in 16b list is `r round(value_16b, 2)` billion dollars, and its ratio to the Chinese claim is `r round(value_16b / 16, 2)`.

```{r 60b, echo=FALSE, cache=FALSE}
lookup_hs_60b <- lookup_hs %>% 
  mutate(tariff = map_chr(hs, find_tariff, df_list_60b))

df2017_60b <- df2017 %>% 
  left_join(lookup_hs_60b, by = "hs")

df_60b <- df2017_60b %>% 
  group_by(tariff) %>% 
  summarize(value = sum(value))

df_60b

value_60b <- df_60b %>% 
  filter(tariff != "character(0)") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

total_value <- value_3b + value_34b + value_16b + value_60b

```
US export value in 60b list is `r round(value_60b, 2)` billion dollars, and its ratio to the Chinese claim is `r round(value_60b / 60, 2)`.

Appearance of strange values like "c(\"16b\", \"16b\")" and "c(\"60b\", \"60b\")" in "tariff" column means that some 8 digit codes fit multiple codes in the look-up table I have prepared. Probably I have omitted too many 0s in the look-up table.

In total, ignoring duplication, China claims it imposed tariffs on 113 billion dollars worth of imports from the US. If my calculation is right, it really is `r round(total_value, 2)` billion dollars, `r round(total_value / 113, 2)` of its claim.

Why are Chinese claims and my calculations so different? My guess is that Chinese calculated values on 6 digit codes by ignoring 8 digit codes. Let us validate it.

## Chinese may have calculated only on 6 digit codes

I cut HTS codes from 8 to 6 digits in the lists of Excel and PDF files.

```{r list_6, echo=FALSE}
df_list_3b_6 <- df_list_3b %>% 
  mutate(
    hs = hs %>% str_c("000000") %>% str_sub(end = 6L)
  ) %>% 
  unique()

df_list_34b_6 <- df_list_34b %>% 
  mutate(
    hs = hs %>% str_c("000000") %>% str_sub(end = 6L)
  ) %>% 
  unique()

df_list_16b_6 <- df_list_16b %>% 
  mutate(
    hs = hs %>% str_c("000000") %>% str_sub(end = 6L)
  ) %>% 
  unique()

df_list_60b_6 <- df_list_60b %>% 
  mutate(
    hs = hs %>% str_c("000000") %>% str_sub(end = 6L)
  ) %>% 
  unique()

```

And I get US export data in HTS 6 digits.

```{r value_6, echo=FALSE}
df2017_6 <- df2017 %>% 
  mutate(
    hs = hs %>% str_c("000000") %>% str_sub(end = 6L)
  ) %>% 
  group_by(hs) %>% 
  summarize(value = sum(value))

value_3b_6 <- df2017_6 %>% 
  semi_join(df_list_3b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000
  
value_34b_6 <- df2017_6 %>% 
  semi_join(df_list_34b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

value_16b_6 <- df2017_6 %>% 
  semi_join(df_list_16b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

value_60b_6 <- df2017_6 %>% 
  semi_join(df_list_60b_6, by = "hs") %>% 
  summarize(value = sum(value)) %>% 
  as.numeric() / 1000000000

```
The results in "3b", "34b", "16b" and "60b" are `r round(value_3b_6, 2)`, `r round(value_34b_6, 2)`, `r round(value_16b_6, 2)` and `r round(value_60b_6, 2)` billion dollars respectively. Ratios to the Chinese claims are `r round(value_3b_6 / 3, 2)`, `r round(value_34b_6 / 34, 2)`, `r round(value_16b_6 / 16, 2)` and `r round(value_60b_6 / 60, 2)`. As Chinese CIF import values are a little bit larger than US FOB export values, and it takes some time to transport goods from US to China, ratios can't be exactly one. These numbers support my guess that Chinese calculated values on 6 digit codes.

After I first finished writing this study, I have found [this page](http://www.customs.go.jp/toukei/sankou/code/code_e.htm).
Among HS 8 digits, the first 6 digits are the same across countries, while the rest digits are domestic, thus can be different across countries. As far as the first 6 digits are the same for US and China, my guess does not change.

But I also have happened to find that as for the first 4 digits of airplanes, US export stats code is "8800", while [China import stats code](http://english.customs.gov.cn/Statics/650f7297-a212-4f0e-a5a8-67422b04e550.html) is "8802". If this kind of interpretation difference is large enough to make the first 6 digits non-comparable between US and China, I can't assume China imports from US on some goods are the same as US exports to China. To completely validate my guess, I have to go for China import stats, but so far I can't find convenient APIs for them.

## Even if retaliation size is less than claimed, US exports are hurt

I get data from 2017 to 2019 based on 6 digit codes, and put tariff category on each item.

```{r get_data_2018_2019, echo=FALSE}
# transform from 8 to 6 digits
df2017m <- df2017 %>% 
    mutate(
      hs = str_sub(hs, end = 6) # cut codes
    ) %>% 
    # summarize because cutting codes may have created duplication
    group_by(time, hs) %>% 
    summarize(value = sum(value)) %>% 
    ungroup() %>% 
  arrange(hs, time)

# get data, takes time
df2018m <- export_to_country(5700, 6, 2018)

df2019m <- export_to_country(5700, 6, 2019)

# combine data
df_m <- df2017m %>% 
  bind_rows(df2018m) %>% 
  bind_rows(df2019m)

hs_category <- df_m %>% 
  select(hs) %>% 
  unique()

hs_category <- hs_category %>% 
  left_join(df_list_3b_6, by = "hs") %>% 
  left_join(df_list_34b_6, by = "hs") %>% 
  left_join(df_list_16b_6, by = "hs") %>% 
  left_join(df_list_60b_6, by = "hs") %>% 
  mutate(double = is.na(tariff.x) + is.na(tariff.y) +
           is.na(tariff.x.x) + is.na(tariff.y.y))

temp <- hs_category %>% 
  filter(double < 3)

```
There are `r nrow(temp)` duplication. I use "first in" method to put category.

```{r hs_category_list, echo=FALSE}
hs_category <- hs_category %>% 
  mutate(
    category = if_else(!is.na(tariff.x), "3b",
                       if_else(!is.na(tariff.y), "34b",
                               if_else(!is.na(tariff.x.x), "16b",
                                       if_else(!is.na(tariff.y.y),
                                               "60b","rest")))
    )
  ) %>% 
  select(-starts_with("tariff"), -double)

```
Retaliation effective dates are April 2, July 6, August 23, September 24, 2018 for "3b", "34b", "16b" and "60b" respectively. "3b", "16b" and "60b" are falling after the effective dates. "34b" is volatile, and is falling from the peak of July-December 2017 before the effective date. "34b" includes agricultural goods like soybeans, automobiles and marine products. Non-retaliated "rest" is not growing. Overall, US exports are hurt.

```{r line_chart, echo=FALSE}
df_m %>% 
  left_join(hs_category, by = "hs") %>% 
  group_by(time, category) %>% 
  summarize(value = sum(value)) %>% 
  ungroup() %>% 
  ggplot(aes(x = time, y = value, color = fct_reorder2(category, time, value))) +
  geom_line() +
    labs(
    title = "US exports to China by each tariff schedule on 6 digits",
    x = "",
    y = "dollars per month",
    color = "category"
  )

```

EOL
