---
title: "Who replaces China in US imports?"
author: "Mitsuo Shiota"
date: "2019-08-06"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Updated: `r Sys.Date()`

## Summary

US imports from China are decreasing in 2019 due to tariffs, and in 2020 due to novel coronavirus. If trade diversion is goin on, other countries are replacing China. In this study, I somewhat arbitrarily choose 4 candidate countries like Vietnam, Korea, Japan and Mexico, and check whether the US imports from these countries increase by HS 4 digit codes.


## Libraries and functions

I use self-made package `tqr`.

```{r libraries, include=FALSE}
# Libraries
library(tidyverse)
library(tsibble)
library(tqr)

# Functions
# get country import data by year and by specified HTS digits (1-10)
import_from_country <- function(country, hs_digits, year) {
  try(if(hs_digits < 1 | hs_digits > 10) stop("hs_digits must be between 1 and 10"))
  
  digits <- if_else(hs_digits <=2, 2,
                    if_else(hs_digits <= 4, 4,
                            if_else(hs_digits <= 6, 6, 10)
                    )
  )
  
  df <- censusapi::getCensus(
    name = "timeseries/intltrade/imports/hs",
    key = keyring::key_get("census"),
    vars = c("CTY_NAME", "GEN_CIF_MO", "I_COMMODITY"),
    time = year,
    CTY_CODE = country,
    COMM_LVL=str_c("HS", digits)
  )
  
  df$GEN_CIF_MO <- as.numeric(df$GEN_CIF_MO)
  
  df$time <- as.Date(paste0(df$time, "-01"), "%Y-%m-%d")
  
  df %>% 
    rename(
      country = CTY_NAME,
      value = GEN_CIF_MO,
      hs = I_COMMODITY
    ) %>% 
    mutate(
      hs = str_sub(hs, end = hs_digits) # cut codes
    ) %>% 
    # summarize because cutting codes may have created duplication
    group_by(country, time, hs) %>% 
    summarize(value = sum(value)) %>% 
    ungroup() %>% 
    select(country, time, hs, value) %>% 
    arrange(country, hs, time)
}

```
## Download data

I download US monthly imports from China, Vietnam, Korea, Japan and Mexico by HS 4 digit codes since 2017, convert to quartely, and fill 0s in gaps.

```{r API_Key_input, include=FALSE}
# Prepare API key
keyring::key_set("census")

```

```{r download_data, echo=FALSE, cache=FALSE}
# Download data
# import from Vietnum and China
# https://www.census.gov/foreign-trade/schedules/c/countryname.html
import <- map_dfr(c(2010, 5520, 5700, 5800, 5880), import_from_country,
                   hs_digits = 4, year = "from 2017 to 2020")

# to million dollars
import <- import %>% 
  mutate(value = value / 1000000)

# from monthly to quarterly
import_q <- import %>% 
  mutate(quarter = yearquarter(time)) %>% 
  group_by(country, hs, quarter) %>% 
  summarize(value = sum(value)) %>% 
  ungroup()

# fill 0
import_q <- import_q %>% 
  spread(key = country, value = value, fill = 0) %>% 
  gather(key = "country", value = "value", -hs, -quarter)

# fill 0
import_q <- import_q %>% 
  spread(key = hs, value = value, fill = 0) %>% 
  gather(key = "hs", value = "value", -quarter, -country)


```

After I fill zeros in gaps, the number of rows is `r nrow(import_q)`. As it matches 5 (number of countries) times `r import_q$hs %>% unique() %>% length()` (number of unique hs goods), I can confirm there are no gaps, and safely use `tqr` package which utilizes lags.


## Choose HS codes of which imports from China declined significantly in 2020 1Q

```{r china_loss, echo=FALSE}
# tsibble
import_q <- import_q %>% 
  as_tsibble(index = quarter, key = c(country, hs))

# goods of whch imports from china declines in 2020 1Q by more than 200 million
# dollars
hs_china_loss <- import_q %>% 
  filter(country == "CHINA") %>% 
  tq_diff(n = 4) %>% 
  as_tsibble() %>% 
  filter(quarter == "2020-01-01", value < -200) %>% 
  `[[`("hs")

# and imports from china in 2020 1Q still more than 500 million dollars
hs_china_loss2 <- import_q %>% 
  filter(hs %in% hs_china_loss) %>% 
  filter(quarter == "2020-01-01", value > 500) %>% 
  `[[`("hs")

```

I first choose HS codes of which imports from China declined by more than 200 million dollars in 2020 1Q from a year ago. Next I choose HS codes of which imports from China are still more than 500 million dollars in 2020 1Q.

## Draw charts of US import changes from a year earlier from each country

For the goods descriptions of HS codes, refer to http://www.kanzei.or.jp/statistical/expstatis/headline/hs2dig/e/85.

Most notable case is that Vietnam replaces China as a main export base of 8517 (telephone sets), probably smart phones. Smart phones are not yet imposed tariffs by the US. This replacement is caused by Samsung's strategy to expand smart phone production capacity in Vietnam, and by ZTE and Huawei losing shares in the US. 

Vietnam is also advancing in 8544 (Cables), 9401 (Seats) and 9403 (Other furniture).

Japan, South Korea and Mexico all supplant China in 8523 (Recording medium).

Mexico replaces China in 8708 (Auto parts).  

```{r draw_charts, echo=FALSE, fig.width=8, fig.height=16}
# draw charts
# http://www.kanzei.or.jp/statistical/expstatis/headline/hs2dig/e/85
import_q %>% 
  filter(hs %in% hs_china_loss2) %>% 
  tq_diff(n = 4) %>% 
  filter(!is.na(value)) %>% 
  mutate(quarter = zoo::as.yearqtr(quarter)) %>% 
  ggplot(aes(x = quarter, y = value, color = country)) +
  zoo::scale_x_yearqtr(format = "%y %qQ") +
  geom_hline(yintercept = 0, size = 2, color = "white") +
  geom_line() +
  facet_wrap(~ hs, scales = "free_y") +
  labs(
    title = "US import changes, quarters, YoY",
    x = "", y = "million dollars", color = ""
  ) +
  theme(legend.position = "bottom")

```

EOL