---
title: "Who replaces China in US imports?"
author: "Mitsuo Shiota"
date: "2019-08-06"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Updated: `r Sys.Date()`

## Summary

US imports from China are decreasing in 2019 due to tariffs, and in 2020 due to novel coronavirus. If trade diversion is goin on, other countries are replacing China. In this study, I somewhat arbitrarily choose 4 candidate countries like Vietnam, Korea, Japan and Mexico, and check whether the US imports from these countries increase by HS 4 digit codes.


## Libraries and functions

I use `tidyverse` and `tsibble` packages.

```{r libraries, include=FALSE}
# Libraries
library(tidyverse)
library(tsibble)


# Functions
# get country import data by year and by specified HTS digits (1-10)
import_from_country <- function(country, hs_digits, year) {
  try(if(hs_digits < 1 | hs_digits > 10) stop("hs_digits must be between 1 and 10"))
  
  digits <- if_else(hs_digits <=2, 2,
                    if_else(hs_digits <= 4, 4,
                            if_else(hs_digits <= 6, 6, 10)
                    )
  )
  
  df <- censusapi::getCensus(
    name = "timeseries/intltrade/imports/hs",
    key = keyring::key_get("census"),
    vars = c("CTY_NAME", "GEN_CIF_MO", "I_COMMODITY"),
    time = year,
    CTY_CODE = country,
    COMM_LVL=str_c("HS", digits)
  )
  
  df$GEN_CIF_MO <- as.numeric(df$GEN_CIF_MO)
  
  df$time <- as.Date(paste0(df$time, "-01"), "%Y-%m-%d")
  
  df %>% 
    rename(
      country = CTY_NAME,
      value = GEN_CIF_MO,
      hs = I_COMMODITY
    ) %>% 
    mutate(
      hs = str_sub(hs, end = hs_digits) # cut codes
    ) %>% 
    # summarize because cutting codes may have created duplication
    group_by(country, time, hs) %>% 
    summarize(value = sum(value), .groups = "drop_last") %>% 
    ungroup() %>% 
    select(country, time, hs, value) %>% 
    arrange(country, hs, time)
}

```
## Download data

I download US monthly imports from China, Vietnam, Korea, Japan and Mexico by HS 4 digit codes since 2017, convert to quartely, and fill 0s in gaps.

```{r API_Key_input, include=FALSE}
# Prepare API key
keyring::key_set("census")

```

```{r download_data, echo=FALSE, cache=FALSE}
# Download data
# https://www.census.gov/foreign-trade/schedules/c/countryname.html
import <- map_dfr(c(2010, 5520, 5700, 5800, 5880), import_from_country,
                   hs_digits = 4, year = "from 2018 to 2020")

# to million dollars
import <- import %>% 
  mutate(value = value / 1000000)

# from monthly to quarterly
import_q <- import %>% 
  mutate(quarter = yearquarter(time)) %>% 
  group_by(country, hs, quarter) %>% 
  summarize(value = sum(value), .groups = "drop_last") %>% 
  ungroup()

# fill 0
import_q <- import_q %>% 
  spread(key = country, value = value, fill = 0) %>% 
  gather(key = "country", value = "value", -hs, -quarter)

# fill 0
import_q <- import_q %>% 
  spread(key = hs, value = value, fill = 0) %>% 
  gather(key = "hs", value = "value", -quarter, -country)


```


## Choose HS codes of which imports from China declined significantly from 2018 Q1

```{r china_loss, echo=FALSE}
# tsibble
import_q <- import_q %>% 
  as_tsibble(index = quarter, key = c(country, hs))

# change value to difference with 2018 Q1
import_q_diff <- import_q %>% 
  group_by(country, hs) %>% 
  mutate(value = value - value[quarter == yearquarter("2018-01-01")]) %>% 
  ungroup()

# goods of whch imports from china declines in 2020 2Q by more than 200 million
# dollars from 2018 Q1
hs_china_loss <- import_q_diff %>% 
  filter(country == "CHINA") %>% 
  filter(quarter == yearquarter("2020-04-01"), value < -200) %>% 
  pull("hs")

# and imports from china in 2020 2Q still more than 500 million dollars
hs_china_loss2 <- import_q %>% 
  filter(hs %in% hs_china_loss) %>% 
  filter(quarter == yearquarter("2020-04-01"), value > 500) %>% 
  pull("hs") %>% 
  unique()

```

I first choose HS codes of which imports from China declined by more than 200 million dollars from 2018 Q1. Next I choose HS codes of which imports from China are still more than 500 million dollars.

## Draw charts of US import changes from 2018 Q1 from each country

For the goods descriptions of HS codes, I refer to https://dataweb.usitc.gov/classification/commodity-description/HTS/4.

```{r hts_4_descriptions, echo=FALSE}
descriptions <- read_csv("data/descriptions.csv", 
    col_names = FALSE, col_types = cols(X1 = col_character()))

names(descriptions) <- c("hs", "description")

descriptions <- descriptions %>% 
  mutate(
    hs = str_c("0", hs),
    hs = str_sub(hs, start = -4L),
    description = str_to_lower(description),
    description = str_wrap(description, width = 32),
    description = str_c(hs, description, sep = "\n")
    )

descri_vector <- descriptions$description
names(descri_vector) <- descriptions$hs

```

Most notable case is that Vietnam replaces China as a main export base of 8517 (telephone sets), probably smart phones. Smart phones are not yet imposed tariffs by the US. This replacement is caused by Samsung's strategy to expand smart phone production capacity in Vietnam, and by ZTE and Huawei losing shares in the US. 

Vietnam is also advancing in 9401 (Seats) and 9403 (Other furniture).

South Korea supplants China in 8523 (Recording medium).

No winners in 8708 (Auto parts).  

```{r draw_charts, echo=FALSE, fig.width=8, fig.height=32}
# draw charts
import_q_diff %>% 
  filter(hs %in% hs_china_loss2) %>% 
  ggplot(aes(x = quarter, y = value, color = country)) +
  geom_hline(yintercept = 0, size = 2, color = "white") +
  geom_line() +
  facet_wrap(~ hs, ncol = 3, scales = "free_y", labeller = labeller(hs = descri_vector)) +
  labs(
    title = "US import changes from 2018 Q1, quarters",
    x = NULL, y = "million dollars", color = NULL
  ) +
  scale_x_yearquarter(date_labels = "%y %q") +
  theme(legend.position = "top")

```

EOL