---
title: "Who pays tariffs?"
author: "Mitsuo Shiota"
date: "2019-05-13"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Updated: `r Sys.Date()`

## Motivation and purpose

On May 11, 2019, Japanese newspaper Nikkei reported that China pays most of the tariffs by cutting export prices, basically confirming [the US President's tweet](https://twitter.com/realDonaldTrump/status/1126815126584266753). I was surprised, as [US import price index: China](https://fred.stlouisfed.org/series/CHNTOT) has not yet declined so much. So I have decided to check how US import prices from China changed from the first half of 2018, just before the first tranche "34b" became effective, for each HTS 10 digit code item, and see the changes by tariff schedule category.

## Libraries and functions

Looking at [census data page](https://api.census.gov/data/timeseries/intltrade/imports/hs/variables.html), I decide to select "GEN_VAL_MO" as value, and "CON_QY1_MO" as quantity, though I selected "GEN_CIF_MO" as value in [the original analysis](README.md).

```{r libraries, include=FALSE}
# Libraries
library(tidyverse)

# Functions
# transform df to 4 column df of time, hs10, hs8 and value
sum2hs8 <- function(df) {
    df$GEN_VAL_MO <- as.numeric(df$GEN_VAL_MO)
    df$CON_QY1_MO <- as.numeric(df$CON_QY1_MO)
    
    df$time <- as.Date(paste0(df$time, "-01"), "%Y-%m-%d")
    
    df %>% 
      rename(
        hs10 = I_COMMODITY,
        value = GEN_VAL_MO,
        quantity = CON_QY1_MO
      ) %>% 
      mutate(
        hs8 = str_sub(hs10, end = -3L) # cut codes from 10 to 8 digits
      ) %>% 
      select(time, hs10, hs8, value, quantity) %>% 
      arrange(time, hs10)
  }

# get country import data by year
import_from_country <- function(country, year) {
  df <- censusapi::getCensus(
    name = "timeseries/intltrade/imports/hs",
    key = keyring::key_get("census"),
    vars = c("I_COMMODITY", "GEN_VAL_MO", "CON_QY1_MO", "UNIT_QY1"),
    time = year,
    CTY_CODE = country,
    COMM_LVL = "HS10"
    )

  df %>% 
    sum2hs8()
}

```

## Extract HTS 8 digit codes for "34b", "16b", "200b" and "300b_a", and 10 digit codes for "excl"

I load tariff lists I have saved in [the separate page](Extract-hts-from-USTR.md).

```{r load_tariff_list, echo=FALSE, results="hide", cache=FALSE}
# load data
load("data/tariff_list.rdata")

df_list_excl <- tibble(
  trf_excl = TRUE,
  hs10 = exclusion_list %>% unlist() %>% unique()
  ) %>% 
  arrange(hs10)

```

## Get monthly import data

```{r API_Key_input, include=FALSE}
keyring::key_set("census")

```

```{r get_data, echo=FALSE, warning=FALSE, cache=FALSE}
df_m_raw <- import_from_country(5700, "from 2018 to 2020")

```

The number of HTS 10 digit code items are `r df_m_raw$hs10 %>% unique() %>% length()`.

## Index unit price of the first half of 2018 as 100

As I have to drop so many data to calculate unit price as value divided by quantity in a specific month, I index unit price as of the first half of 2018 equals to 100.

```{r index_unit_price, echo=FALSE}
unit_price_base <- df_m_raw %>%
  filter(time >= "2018-01-01", time <= "2018-06-01") %>% 
  group_by(hs10) %>% 
  summarize(
    value = sum(value),
    quantity = sum(quantity),
    .groups = "drop_last"
  ) %>% 
  mutate(unit_price = value / quantity) %>% 
  select(hs10, unit_price)

df_m <- df_m_raw %>% 
  mutate(unit_price = value / quantity) %>% 
  inner_join(unit_price_base, by = "hs10") %>% 
  mutate(index = unit_price.x / unit_price.y * 100) %>% 
  drop_na(index) %>% 
  filter(index < Inf)

```

I add tariff schedule category ("34b", "16b", "200b", "300b_a", "rest") to each item.

```{r add_category, echo=FALSE}
# add category
df_m <- df_m %>% 
  left_join(df_list_34b, by = "hs8") %>% 
  left_join(df_list_16b, by = "hs8") %>% 
  left_join(df_list_200b, by = "hs8") %>% 
  left_join(df_list_300b_a, by = "hs8") %>% 
#  left_join(df_list_300b_c, by = "hs8") %>% 
  left_join(df_list_excl, by = "hs10") %>% 
  replace_na(list(
    trf_34b = FALSE,
    trf_16b = FALSE,
    trf_200b = FALSE,
    trf_300b_a = FALSE,
#    trf_300b_c = FALSE,
    trf_excl = FALSE
  )) %>% 
  mutate(
    category_old = if_else(trf_34b, "34b",
                       if_else(trf_16b, "16b",
                               if_else(trf_200b, "200b",
                                       if_else(trf_300b_a, "300b_a", "rest")))),
    category = if_else(trf_excl, "excl", category_old)
  ) %>% 
  select(-starts_with("trf_"), -category_old)

df_m$category <- factor(df_m$category, 
                            levels = c("34b", "16b", "200b", "300b_a", "excl", "rest"))

```


The number of HTS 10 digit code items by category are as below.

```{r n_by_category, echo=FALSE}
df_m %>% 
  select(hs10, category) %>% 
  unique() %>% 
  group_by(category) %>% 
  count()

```


## Look at the price changes from the first half of 2018

As I index the first half of 2018 as 100, the price range in this period is naturally narrow. “34b”, “16b”, “200b” and "300b_a" imposed tariffs got effective on July 6, 2018, August 23, 2018, September 24, 2018, and September 1, 2019, respectively. If Chinese pay tariffs, prices fall after the effective date.

```{r boxplot, echo=FALSE, fig.width=8, fig.height=6}
df_m %>% 
  ggplot(aes(time, index)) +
  geom_boxplot(aes(group = time)) +
  facet_wrap(~ category) +
  coord_cartesian(ylim = c(0, 200)) +
  labs(
    title = "Price index",
    x ="", y = "index (2018 1H = 100)"
  )

```

In some non-commodity items, products are so different, and it is difficult to interpret value per quantity changes as price changes. So I think we had better ignore outliers, and look at the medians of price indices by each category.

```{r line_chart, echo=FALSE, warning=FALSE, fig.width=6, fig.height=6}
df_sum <- df_m %>% 
  group_by(category, time) %>% 
  summarize(index = median(index), .groups = "drop_last")

df_sum %>% 
  ggplot(aes(time, index,
             color = fct_reorder2(category, time, index))) +
  geom_hline(yintercept = 100, color = "white", size = 2) +
  geom_line() +
  labs(
    title = "Median price index by category",
    x = "", y = "Median price index (2018 1H = 100)", color = ""
  )

```

In the latest month, median price indices by category are as below.

```{r price_decline, echo=FALSE}
df_sum2 <- df_sum %>% 
  filter(time == max(df_sum$time))

df_sum2

```

Chinese are paying `r round(100 - as.numeric(df_sum2[1, 3]), 1)` out of 25 percent in "34b", `r round(100 - as.numeric(df_sum2[2, 3]), 1)` out of 25 in "16b", `r round(100 - as.numeric(df_sum2[3, 3]), 1)` out of 25 in "200b", and `r round(100 - as.numeric(df_sum2[4, 3]), 1)` out of 7.5 in "300b_a" in the latest month. Should I subtract `r round(100 - as.numeric(df_sum2[6, 3]), 1)` percent decline of "rest"?

## Correction

On October 26, 2020, I found price index movements above are not compatible with [US import price index: China, in total](https://fred.stlouisfed.org/series/CHNTOT). In 2020, every price index of tariff schedule category above shows rapid decline, while the total price index shows stabilization or even a bit of increase.

Then I noticed there are a lot of zeroes in 2020 in the price index box plots above, especially those of "34b" and "excl". Price index, which is value divided by quantity and is standardized, can't be zero. It turned out there are `r df_m %>% filter(index == 0) %>% nrow()` zero indices.

Where these zero indices come from? Small parts, `r df_m %>% filter(value == 0) %>% nrow()`, come from zero values. Most parts must come from standardization. It leads to my guess that unit is revised at the start of every year.

```{r zero_price}
# zero indices in 2018
df_m %>% 
  filter(index == 0, time >= "2018-01-01", time < "2019-01-01") %>% 
  nrow()

# in 2019
df_m %>% 
  filter(index == 0, time >= "2019-01-01", time < "2020-01-01") %>% 
  nrow()

# in 2020
df_m %>% 
  filter(index == 0, time >= "2020-01-01") %>% 
  nrow()

```

If unit is inconsistent over years, my method of calculation of price index is meaningless.

Anyway, I remove zero value and zero index, and recalculate below.

```{r recalculate, echo=FALSE}
df_m_no_zero_value <- df_m_raw %>% 
  filter(value != 0)

unit_price_base <- df_m_no_zero_value %>%
  filter(time >= "2018-01-01", time <= "2018-06-01") %>% 
  group_by(hs10) %>% 
  summarize(
    value = sum(value),
    quantity = sum(quantity),
    .groups = "drop_last"
  ) %>% 
  mutate(unit_price = value / quantity) %>% 
  select(hs10, unit_price)

df_m2 <- df_m_no_zero_value %>% 
  mutate(unit_price = value / quantity) %>% 
  inner_join(unit_price_base, by = "hs10") %>% 
  mutate(index = unit_price.x / unit_price.y * 100) %>% 
  drop_na(index) %>% 
  filter(index < Inf, index > 0)

# add category
df_m2 <- df_m2 %>% 
  left_join(df_list_34b, by = "hs8") %>% 
  left_join(df_list_16b, by = "hs8") %>% 
  left_join(df_list_200b, by = "hs8") %>% 
  left_join(df_list_300b_a, by = "hs8") %>% 
#  left_join(df_list_300b_c, by = "hs8") %>% 
  left_join(df_list_excl, by = "hs10") %>% 
  replace_na(list(
    trf_34b = FALSE,
    trf_16b = FALSE,
    trf_200b = FALSE,
    trf_300b_a = FALSE,
#    trf_300b_c = FALSE,
    trf_excl = FALSE
  )) %>% 
  mutate(
    category_old = if_else(trf_34b, "34b",
                       if_else(trf_16b, "16b",
                               if_else(trf_200b, "200b",
                                       if_else(trf_300b_a, "300b_a", "rest")))),
    category = if_else(trf_excl, "excl", category_old)
  ) %>% 
  select(-starts_with("trf_"), -category_old)

df_m2$category <- factor(df_m2$category, 
                            levels = c("34b", "16b", "200b", "300b_a", "excl", "rest"))

```


The number of HTS 10 digit code items by category are as below.

```{r n_by_category2, echo=FALSE}
df_m2 %>% 
  select(hs10, category) %>% 
  unique() %>% 
  group_by(category) %>% 
  count()

```

Box plots.

```{r boxplot2, echo=FALSE, fig.width=8, fig.height=6}
df_m2 %>% 
  ggplot(aes(time, index)) +
  geom_boxplot(aes(group = time)) +
  facet_wrap(~ category) +
  coord_cartesian(ylim = c(0, 200)) +
  labs(
    title = "Price index",
    x ="", y = "index (2018 1H = 100)"
  )

```

Line charts.

```{r line_chart2, echo=FALSE, warning=FALSE, fig.width=6, fig.height=6}
df_sum <- df_m2 %>% 
  group_by(category, time) %>% 
  summarize(index = median(index), .groups = "drop_last")

df_sum %>% 
  ggplot(aes(time, index,
             color = fct_reorder2(category, time, index))) +
  geom_hline(yintercept = 100, color = "white", size = 2) +
  geom_line() +
  labs(
    title = "Median price index by category",
    x = "", y = "Median price index (2018 1H = 100)", color = ""
  )

```

In the latest month, median price indices by category are as below.

```{r price_decline2, echo=FALSE}
df_sum2 <- df_sum %>% 
  filter(time == max(df_sum$time))

df_sum2

```

Chinese are paying `r round(100 - as.numeric(df_sum2[1, 3]), 1)` out of 25 percent in "34b", `r round(100 - as.numeric(df_sum2[2, 3]), 1)` out of 25 in "16b", `r round(100 - as.numeric(df_sum2[3, 3]), 1)` out of 25 in "200b", and `r round(100 - as.numeric(df_sum2[4, 3]), 1)` out of 7.5 in "300b_a" in the latest month. Should I subtract `r round(100 - as.numeric(df_sum2[6, 3]), 1)` percent decline of "rest"?

Looks less incompatible with [US import price index: China, in total](https://fred.stlouisfed.org/series/CHNTOT), though volatile "34b" suggests I had better cut off high indices at certain level.

EOL